\textsc{Computing Summands of the Frobenius \\ over Multigraded Cox Rings}

### 2024-07-28 beamer/JMM/beamer.tex
moved JMM beamer files

  %\pause
  - Hartshorne: any $L\in\Pic\PPn$, $F_*L$ splits as a sum of line bundles.
  %\pause
  \begin{block}{Example: $F_*\o_{\PPn}$}
    \vspace*{-0.2in}
    \begin{align*}
      \text{When $p=3$, $n=2$:} \\
      F_*\o_{\PP2} = \o \oplus &\o(-1)^7 \oplus \o(-2). \\
      \text{When $p=2$, $n=5$:} \\
      F_*\o_{\PP5} = \o \oplus &\o(-1)^{15} \oplus \o(-2)^{15} \oplus \o(-3). \\
      F_*^2\o_{\PP5} = \o \oplus &\o(-1)^{120} \oplus \o(-2)^{546} \oplus \o(-3)^{336} \oplus \o(-4)^{21}
    \end{align*}
  \end{block}
\end{frame}

@@

  %\pause
  - Bogvad '98, Thomsen '00: $F_*L$ totally splits for any $L\in\Pic X$.
  %\pause
  \begin{block}{Example: $X = $ Hirzebruch surface of type 3, $F_*\o_X, p = 2$}
    \vspace*{-0.2in}
    \[ F_*\o_X = \o_X \oplus \o_X(-1,0)^2 \oplus \o_X(0,-1)^2 \oplus \o_X(1,-1)^3 \oplus \o_X(2,-1). \]
  \end{block}

@@

\begin{frame}{Frobenius pushforward on Mori Dream Spaces}
  \begin{block}{Example: $X = \mathrm{Bl}_4\PP2, F_*\o_X, p = 2$}
    \vspace*{-0.15in}
    h\begin{align*}
      F_*^2\o_X = {\o_{X}^{1}}
      &\oplus {\o_{X}^{2}\ \left(-2,\,1,\,1,\,1,\,1\right)} \oplus {\o_{X}^{2}\ \left(-1,\,0,\,0,\,0,\,1\right)} \\ 
      &\oplus {\o_{X}^{2}\ \left(-1,\,0,\,0,\,1,\,0\right)} \oplus {\o_{X}^{2}\ \left(-1,\,0,\,1,\,0,\,0\right)} \\
      &\oplus {\o_{X}^{2}\ \left(-1,\,1,\,0,\,0,\,0\right)} \oplus B \oplus G,
    \end{align*}
    where $B, G$ are rank 3 and rank 2 indecomposables. \\
    \hfill (c.f. [Hara, '15])
  \end{block}
\end{frame}

\begin{frame}{The splitting algorithm}
  \begin{alertblock}{The ``Meataxe'' algorithm}
    Input: finitely generated graded module $M$ over an algebra.
    %\pause
    \begin{enumerate}
    \item compute $\Hom(M, M)$; %\pause
    \item find element $h$ corresponding to an idempotent; %\pause
    \item find the image and cokernel of $h$; %\pause
    \item repeat.
    \end{enumerate}
  \end{alertblock}
\end{frame}

>>>>

  %\pause
  - Hartshorne: any $L\in\Pic\PPn$, $F_*L$ splits as a sum of line bundles. \\
  %\pause
  - Raczka '24: any v.b.\ on $\PPn$, $F_*^e\cE$ splits as a sum of $\OO$'s and $\Omega$'s.
  %\pause
  \begin{block}{Example: $F_*\o_{\PPn}$}
    \vspace*{-0.2in}
    \begin{align*}
      \text{When $p=3$, $n=2$:} \\
      F_*\o_{\PP2} = \o \oplus &\o(-1)^7 \oplus \o(-2). \\
      \text{When $p=2$, $n=5$:} \\
      F_*\o_{\PP5} = \o \oplus &\o(-1)^{15} \oplus \o(-2)^{15} \oplus \o(-3). \\
      F_*^2\o_{\PP5} = \o \oplus &\o(-1)^{120} \oplus \o(-2)^{546} \oplus \o(-3)^{336} \oplus \o(-4)^{21}
    \end{align*}
  \end{block}
\end{frame}

@@

  %\pause
  - Bogvad '98, Thomsen '00: $F_*L$ totally splits for any $L\in\Pic X$.
  %\pause
  \begin{block}{Example: $X = $ Hirzebruch surface of type 3, $F_*\o_X, p = 3$}
    \vspace*{-0.2in}
    \[ F_*\o_X = \o_X \oplus \o_X(-1,0)^2 \oplus \o_X(0,-1)^2 \oplus \o_X(1,-1)^3 \oplus \o_X(2,-1). \]
  \end{block}

@@

\begin{frame}{Frobenius pushforward on Mori Dream Spaces}
  \begin{block}{Example: $X = \mathrm{Bl}_4\PP2, F_*\o_X, p = 2$}
    \vspace*{-0.15in}
    \begin{align*}
      F_*^2\o_X = {\o_{X}^{1}}
      &\oplus {\o_{X}^{2}\ \left(-2,\,1,\,1,\,1,\,1\right)} \oplus {\o_{X}^{2}\ \left(-1,\,0,\,0,\,0,\,1\right)} \\ 
      &\oplus {\o_{X}^{2}\ \left(-1,\,0,\,0,\,1,\,0\right)} \oplus {\o_{X}^{2}\ \left(-1,\,0,\,1,\,0,\,0\right)} \\
      &\oplus {\o_{X}^{2}\ \left(-1,\,1,\,0,\,0,\,0\right)} \oplus B \oplus G,
    \end{align*}
    where $B, G$ are rank 3 and rank 2 indecomposables. \\
    \hfill (c.f. [Hara, '15])
  \end{block}
\end{frame}

\begin{frame}{The splitting algorithm}
  \begin{alertblock}{The ``Meataxe'' algorithm}
    Input: finitely generated graded module $M$ over an algebra.
    %\pause
    \begin{enumerate}
    \item Compute $\Hom(M, M)$; \\ %\pause
      Note: the degree zero part only, so \textbf{multigrading helps!} %\pause
    \item Find element $h$ corresponding to an idempotent; \\ %\pause
      \begin{itemize}
      \item Hope that generators of $\Hom$ are idempotents; %\pause
      \item In finite characteristic, use a general endomorphism. %\pause
      \end{itemize}
    \item Find the image and cokernel of $h$;
    \item Repeat.
    \end{enumerate}
  \end{alertblock}

  \vfill
  %\pause
  \begin{lemma}
    Let $M$ be an $R$-module, and let $A\in\End_R(M)$. If the induced action of $A$ on $M/mM$ is idempotent, then $M$ admits a direct sum decomposition $N\oplus M/N$, where $N\cong \im A$.
    % Let $A$ be an endomorphism of a $k$-vector space such that all eigenvalues of $A$ are contained in $k$. If $\lambda$ is an eigenvalue of $A$, then some power of $A-\lambda I$ is an idempotent.
  \end{lemma}

\end{frame}
